---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
  word_document: default
---



### Necessary packages
```{r}
library(tidyverse)
library(readxl)
```


### Working directory
```{r}
getwd()
list.files()
```

###  data import
```{r}
housing_price <- read_excel("C:/Users/Hp/Downloads/housing.xlsx")

head(housing_price)
```

### Data Structure
```{r}
glimpse(housing_price)
```

### Summary statitsica and check for  the missing values (NA’s)  of each column:
```{r}
summary(housing_price)
```


### Check for outliers for the target variables:

#### Outlier detection with boxplot
```{r}
ggplot( data = housing_price, aes(y=median_house_value ))+
  geom_boxplot(outlier.colour = "red")
```

# IQR method
```{r}
summary(housing_price$median_house_value)
```

```{r}
Q1 <-119600
Q3 <- 264725
IQR <- Q3 - Q1
IQR
IQR(housing_price$median_house_value)
```

### filter the outliers from the data 
```{r}
outliers<-  housing_price %>% 
  filter( median_house_value < Q1 - 1.5 *IQR  | median_house_value > Q3 + 1.5 *IQR )

outliers

nrow(outliers)/ nrow(housing_price)*100
```

### Look for the distribution of the target variable Median House Value
```{r}
ggplot( data = housing_price, aes(x=median_house_value ))+
   geom_histogram(color= "white")
  
```

### Transforming the distribution of the Median House Value using with log transformation

### with log transformation
```{r}
ggplot(data = housing_price, aes(x= log(median_house_value )))+
   geom_histogram(color= "white")
   
```



### Detect outliers atr log transformation

## boxplot
```{r}
ggplot(data = housing_price, aes(y= log(median_house_value )))+
   #geom_histogram(color= "white")
   geom_boxplot(outlier.colour = "red")
```

### Applied the  log transfomation to the  median_house_value (the target variable)
```{r}
housing_priceT <- housing_price %>% 
  mutate(log_median_house_value = log(median_house_value))

```


```{r}
summary(housing_priceT)
```

# Applying the IQR method to detect outliers
```{r}
summary(housing_priceT$log_median_house_value)
```

### IQR
```{r}
Q1 <-11.691
Q3 <- 12.486 
IQR <- Q3 - Q1
IQR

```


#outlier detection

```{r}
outliers <- housing_priceT %>% 
  filter( log_median_house_value < Q1 - 1.5 *IQR  | log_median_house_value  > Q3 + 1.5 *IQR )


dim(outliers)

```

we detect 26 points as outliers


### drop outliers
```{r}

housing_priceT <- housing_priceT %>% 
  filter( log_median_house_value > Q1 - 1.5 *IQR  | log_median_house_value < Q3 + 1.5 *IQR )
housing_price
```




### Look for the distribution of the features numerical columns

```{r fig.height=4, fig.width=8}

housing_price %>% 
  pivot_longer(cols =housing_median_age : median_income,names_to = "variables", values_to = "values" ) %>% 
                    
ggplot( aes(x= values)) +
  geom_histogram() +
  facet_wrap(~ variables, scales = 'free_x') 
 
```

### Analyze the correlation of each variable with the target variable

```{r fig.height=4, fig.width=10 }
housing_price %>% 
  pivot_longer(cols =housing_median_age : median_income,names_to = "variables", values_to = "values" ) %>% 
                    
ggplot( aes(x= values, y= median_house_value)) +
  geom_point() +
  facet_wrap(~ variables, scales = 'free_x') 
 
```


###  Correlation Matrix
```{r fig.height=6, fig.width=10}

M<-housing_price %>% 
  drop_na(total_bedrooms) %>% 
   select(-ocean_proximity, ) %>%  cor()
M
```

### Correlation Matrix plot
```{r fig.height=5, fig.width=10}
library(corrplot)
corrplot( M, method = "number")

```
  
  
### Impact of Ocean Proximity on housing price

```{r fig.height=5, fig.width= 8}
ggplot(data= housing_price, aes(x= ocean_proximity, y =median_house_value, fill= ocean_proximity))+
  geom_boxplot()
```


# Plotting longitude and latitude to visualize housing locations

```{r fig.height=5, fig.width= 8}
ggplot(housing_price, aes(x = longitude, y = latitude, color = median_house_value)) +
  geom_point()+
  scale_color_gradient(low = "yellow", high = "red")

```


## Data Cleaning

### Convert ocean_proximity variable to factor
```{r}
 housing_price <- housing_price %>% 
  mutate(ocean_proximity = as.factor(ocean_proximity))

housing_price
```


### Ratio of NA's to al data
```{r}
(sum(is.na(housing_price$total_bedrooms))/ nrow(housing_price)) * 100
```
(about 1%)

### Drop NA's
```{r}
housing_price <- housing_price %>% 
               drop_na(total_bedrooms)
summary(housing_price$total_bedrooms)
```

### Applied the  log transfomation to the  median_house_value (the target variable)
```{r}
housing_price <- housing_price %>% 
  mutate(log_median_house_value = log(median_house_value))

housing_price
```

### Drop the outliers in the target vriable ( median house value).
```{r}
housing_price <- housing_price %>% 
  filter(!median_house_value %in% outliers$log_median_house_value)

```

## last check for aoutliers 
```{r}
ggplot(data = housing_price, aes(y= log_median_house_value ))+
   geom_boxplot(outlier.colour = "red")

```



## Data Preprocessing 
### Scaling or standarzing  the numerical features using the (Min-Max method):

min-max 

z = x- min(x)/ max(x)-mix(x)

normalisation (Scaling)

z = x -mean(x) / sd(x) --> scale

#### The Normalize function
```{r}

normalize <- function(x) {
  result <- (x - min(x, na.rm = TRUE)
  ) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
  
  return(result)
}

```


### apply to the numerical features

```{r}
housing_price <- housing_price %>% 
 mutate(housing_median_age = normalize(housing_median_age),
        total_rooms = normalize(total_rooms), 
        total_bedrooms = normalize(total_bedrooms),
        population = normalize(population),
        households = normalize(households),
        median_income = normalize(median_income)
        )
housing_price
```


### Convert data to factors Dummy variable (one hote encoding)
```{r}
levels(housing_price$ocean_proximity)
```

```{r}
housing_price <- housing_price %>% 
  mutate( `<1H OCEAN` = if_else(ocean_proximity =="<1H OCEAN", 1, 0),
          INLAND = if_else(ocean_proximity =="INLAND", 1, 0),
          ISLAND  = if_else(ocean_proximity =="ISLAND", 1, 0),
          `NEAR BAY` = if_else(ocean_proximity =="NEAR BAY", 1, 0),
          `NEAR OCEAN` = if_else(ocean_proximity =="<1H OCEAN", 1, 0))
housing_price
```

### Final data
```{r}
housing_price <- housing_price %>% 
  select(-median_house_value,-ocean_proximity) 
housing_price
```

## Selecting and Training Machine Learning Models.


### Build a linear regression model using Train/Test -Split

#### Split the dataset into training and testing sets (80-20 ratio)
```{r}
library(rsample)
set.seed(222)
# Save the split information for an 80/20 split of the data
housing_split <- initial_split(housing_price, prop = 0.80)
housing_split

```

#### Split the dataset into training and testing 
```{r}
housing_train <- training(housing_split)
housing_test  <-  testing(housing_split)

dim(housing_train)
dim(housing_test)
```


#### Create a linear regression model using the training dataset
```{r}
lm_model <- lm(formula = log_median_house_value  ~., data =housing_train)
```

### model output
```{r}
summary(lm_model)
```


### Model Evaluation

####  Make predictions on the test dataset using the trained model
```{r}
predictions <- predict(lm_model, newdata = housing_test)
```

2. Evaluate the model's performance using appropriate metrics (e.g., RMSE, MSE, R-squared)


## testing MSE
$$MSE_{Test} = 1/n \sum_{i=1}^{n}(y - \hat{y})^2$$


```{r}
MSE <- mean((housing_test$log_median_house_value - predictions)^2)
MSE

```


### Predict the price for an new House in California

#### Define the new house charactristics
```{r}
+++ b
new_data <- data.frame(longitude =-120.80	,latitude = 37.50 , housing_median_age=  0.41, 
           total_rooms = 0.05,  total_bedrooms = 0.08 ,    
           population = 0.17, households= 0.24, median_income= 0.35,
           `<1H OCEAN`= 0, INLAND = 1,ISLAND = 0,`NEAR BAY` = 0,`NEAR OCEAN` =0
           ) %>%  rename(`<1H OCEAN` = X.1H.OCEAN, 
                         `NEAR OCEAN`= NEAR.OCEAN,
                         `NEAR BAY` = NEAR.BAY )
new_data
```

### Predict the price  using our trainde model
```{r}
predict(lm_model, newdata = new_data)
```
```{r}
exp(11.49573 )
```

